---
title: "2021-03-30 Makeup Shades"
output: github_document
---

```{r setup, include=FALSE}
.libPaths("C:/R-packages2/")
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```



```{r}
library(tidyverse)
library(patchwork)
```


### Load data

```{r}
allCategories <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allCategories.csv')
```
## Building a random name and color generator for new make-up products

This script extracts names from existing products and randomly recombines them, a random hex code (balanced sample for lightness) is then added to the new name and a plot of 9 different invented make-up products is created. 

### Filter for those products that have a double name that can be recombined

```{r}
double_name <- allCategories %>% 
  filter(!is.na(name), 
         str_detect(categories, "location", negate = TRUE),
         str_detect(name, "-", negate = TRUE)) %>% 
  mutate(names = str_split(name, " "), 
         length = lengths(names)) %>% 
  filter(length == 2) 
  
```

### Get sample of products with even distribution of lightness

```{r}
hist(double_name$lightness)

set.seed(123)

sample_even <- double_name %>% 
  group_by(lightness) %>% 
  slice_sample(n= 1) 

sample_even %>% 
  ggplot(aes(lightness)) +
  geom_histogram()
```

### Extract first and last name

```{r}

first <- double_name %>% 
  select(names) %>% 
  unnest(cols = c(names)) %>% 
  slice_head(n = nrow(double_name))
first <- first$names

second <- double_name %>% 
  select(names) %>% 
  unnest(cols = c(names)) %>% 
  slice_tail(n = nrow(double_name))
second <- second$names
```

### Extract HEX

```{r}
hex_code_list <- double_name$hex
```

### Random Make-up generator

```{r}
invent_makeup <- function(first_names = first, 
                          second_names = second,
                          hex_codes= hex_code_list){
  first <- sample(first_names, 1)
  second <- sample(second_names,1)
  hex <- sample(hex_codes, 1)
  out <- data.frame(name = str_to_title(str_c(first, second, sep = " ")),
                    hex = hex)
  return(out)
}

invent_makeup()
```

### Plot new make-up

```{r}

plot_random_makeup <- function(){
  
  make_up <- invent_makeup()
  p <- ggplot(data = make_up, aes(x = 1, y = 1)) +
    theme(plot.background = element_rect(fill = make_up$hex, color = make_up$hex),
          panel.background = element_rect(fill = make_up$hex, color = make_up$hex),
          panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          axis.ticks.x=element_blank(),
          axis.title.x = element_blank(),
          axis.text.x=element_blank()) +
    scale_y_continuous(limits = c(0,1)) +
    scale_x_continuous(limits = c(0,1)) +
    annotate(geom = "text", x = 0.5, y =0.5, 
             label = make_up$name)
  return(p)
}
```

### Combine plots into panel

```{r}
p1 <- plot_random_makeup()
p2 <- plot_random_makeup()
p3 <- plot_random_makeup()
p4 <- plot_random_makeup()
p5 <- plot_random_makeup()
p6 <- plot_random_makeup()
p7 <- plot_random_makeup()
p8 <- plot_random_makeup() 
p9 <- plot_random_makeup() 


p <- p1 + p2 + p3 + p4+ p5 + p6 + p7 + p8 + p9 + plot_annotation(title = "Random Make-up Generator", caption = "Source: The Pudding data | Graphic: @TannerFlorian") &
  theme(plot.background = element_rect(fill = "black", color = "black"),
        text = element_text(color = "white"))

p
```

### Save graph

```{r}

ggsave(p, filename = "Random make-up.png", units = "in", width = 4, height = 2.25, dpi = 300, scale = 1.4)

```



```{r}
sessionInfo()
```


