---
title: "2021-05-25 Mario Kart World Records"
output: github_document
---

```{r setup, include=FALSE}
.libPaths("C:/R-packages2/")
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```



```{r}
library(tidyverse)
library(extrafont)
library(showtext)
library(ggsci)
library(ggimage)
```

## Read data 

```{r}
ds <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/records.csv')
drivers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/drivers.csv')
```



## Visualization

* All times are PAL times
* Most competitive tracks?
* Number of records
* Which tracks seem like they could improve more? 
* Normalization of records?
* Percent improvement on original record? 


```{r}
ds %>% 
  filter(shortcut == "No", type == "Single Lap") %>% 
  ggplot(aes(x= date, y = time, group = track)) +
  geom_line() +
  facet_wrap(~track, scales = "free_y")
```

```{r}
ds %>% 
  filter(date > "2010-01-01", shortcut == "No", type == "Single Lap") %>% 
  ggplot(aes(x= date, y = time, group = track)) +
  geom_line() +
  facet_wrap(~track, scales = "free_y")
```
## Compare change in time to first record

* This graph is no better than the straight times
```{r}
ds %>% 
  filter(date > "1999-01-01", shortcut == "No", type == "Single Lap") %>% 
  group_by(track) %>% 
  mutate(original_record = max(time),
         change_to_original = 1- time/original_record) %>% 
  ggplot(aes(date, change_to_original, group = track)) +
  geom_line() +
  facet_wrap(~track)
```

## Which ones are most competitive?

* Not much difference 
```{r}
ds %>% 
  filter(shortcut == "No", type == "Single Lap") %>% 
  group_by(track) %>% 
  summarise(records = n())
```

## Full graph

* Focus on Bowser's castle
* The playerbase seems relatively small, highlight the race between the different players


### Read background image and symbols

```{r}
bg <- "Capture75opCropmoved.png"
flag <- "flags_white_bg.png"
font_add(family = "mariokart", regular = "SuperMario256.ttf")
showtext_auto()
```

### Define theme

```{r}
mk_font <- "mariokart"

mk64_theme <- theme_dark() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_line(),
        axis.title.x = element_blank(),
        plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black"),
        text=element_text(family=mk_font, color = "white"),
        axis.text = element_text(family=mk_font, color = "white", size = 18),
        plot.title = element_text(family=mk_font, color = "white", size = 33),
        plot.subtitle = element_text(family=mk_font, color = "white", size = 22),
        axis.title.y = element_text(family=mk_font, color = "white", size = 19),
        legend.position = "none")
```


```{r}
p <- ds %>% 
  filter(date > "2005-02-01", shortcut == "No", 
         type == "Single Lap", track == "Bowser's Castle") %>% 
  ggplot(aes(x= date, y = time, group = track)) +
  geom_line(color = "white", size = 2.5) +
  geom_point(aes(fill = player), size = 1.8, shape =21) +
  labs(title = "The race for the world record", y = "Seconds", 
       subtitle = "Four players have been fighting it out since 2005",
       caption = "") +
  scale_fill_jco() +
  scale_x_date(limits = c(as.Date("2005-01-01"), as.Date("2023-01-01"))) + 
  geom_bgimage(bg) +
  geom_image(aes(x= as.Date("2022-10-01"), y = 43.13, image = flag), size = 0.13) +
  # Dan
  annotate(geom = "text", x =  as.Date("2021-02-02"), 
           y = 43.24, label = "Current record:",
           color = "white", family = mk_font, size = 6) +
  annotate(geom = "text", x =  as.Date("2021-02-02"), 
           y = 43.22, label = "'Dan': 43.15 sec",
           color = "white", family = mk_font, size = 6) +
  annotate(geom = "segment", x = as.Date("2021-02-02"), xend = as.Date("2021-02-02") , 
           y = 43.21, yend = 43.16, color = "white", size = 0.65) +
  # Abney317
  annotate(geom = "text", x =  as.Date("2015-01-01"), 
           y = 43.17, label = "'Abney317'",
           color = "white", family = mk_font, size = 6) +
  annotate(geom = "segment", x = as.Date("2016-06-01"), xend = as.Date("2018-05-01"), 
           y = 43.17, yend = 43.185, color = "white", size = 0.65) +
  # MJ
  annotate(geom = "text", x =  as.Date("2005-04-01"), 
           y = 43.32, label = "'MJ'",
           color = "white", family = mk_font, size = 6) +
  # MR
  annotate(geom = "text", x =  as.Date("2013-01-01"), 
           y = 43.30, label = "'MR'",
           color = "white", family = mk_font, size = 6) +
  mk64_theme + 
  labs(caption = "Source: Benedikt Claus/mkwrs.com | Visualization: @TannerFlorian")
p
```

```{r}
ggsave("mario.png", p, device = "png", units = "in", width = 4, height = 2.25, scale = 1, type = "cairo")
```


# https://www.dafont.com/super-mario-256.font

```{r}
sessionInfo()
```


