---
title: "2021-10-26 Ultra Trail Running"
output: github_document
---

```{r setup, include=FALSE}
.libPaths("C:/R-packages2/")
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r, message= FALSE, marning = FALSE, echo = FALSE}
library(tidyverse)
library(lubridate)
library(showtext)
library(ggtext)
# sysfonts::font_add_google("Roboto Condensed")
`%nin%` <- Negate(`%in%`)
```

## Read data 

```{r}
ultra_rankings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-26/ultra_rankings.csv')
race <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-26/race.csv')

```

## Ideas 

*	Rank races by difficulty
*	Average finishing times
*	Lengths, vert
*	Competitiveness
*	Map of events with initial year?
  * Not really enough data + COVID effect 
*	How much slower is average than the winners?
*	Big wins, course records? (% faster than second place)
*	Cut-offs? 
* DNF - rate

## EDA

* 0 participants in 'race' df is wrong


```{r}
event_creation <- race %>% 
  filter(participants > 0) %>% 
  group_by(race_year_id, race) %>% 
  summarise(first_date = min(date)) %>% 
  mutate(first_year = floor_date(first_date, unit = "year")) %>% 
  ungroup()
```

```{r}
event_creation %>% 
  count(first_year)
```

```{r}
race %>%
  filter(participants > 0) %>% 
  mutate(year = floor_date(date, unit = "year")) |>
  group_by(year) |> 
  count()
```



* Long-standing events

```{r}
race %>% 
  filter(participants > 0,
         distance  > 0) %>% 
  group_by(event, race) %>% 
  count() %>% 
  arrange(desc(n))
```

```{r}
hist(race$distance)
```


* None have 0 participants
* Naming is very inconsistent


```{r}
ultra_rankings |>
  left_join(race) |>
  group_by(race_year_id, race, date, event) |>
  count() |>
  arrange(desc(n))
```


### Pace of 100-mile events

* Using 2018 data for simplicity, many iconic races are included 

```{r}
races_2018 <- ultra_rankings |>
  left_join(race) |>
  filter(distance > 155, 
         distance < 170,
         date > as.Date("2017-12-31"),
         date < as.Date("2018-12-31")) |>
  mutate(miles = distance / 1.60934,
         pace_km = (time_in_seconds/60)/distance,
         pace_mile = (time_in_seconds/60)/miles) |>
  group_by(race_year_id) |>
  summarise(avg_pace_mile = mean(pace_mile, na.rm = TRUE),
            avg_pace_km = mean(pace_km, na.rm = TRUE),
            race_year_id = race_year_id,
            fastest_pace_km = min(pace_km, na.rm = TRUE),
            participants_from_rankings = n()) |>
  left_join(race) |>
  distinct() |>
  filter(participants_from_rankings > 50) %>% 
  ungroup()
```

```{r}
skimr::skim(races_2018)
```


```{r}
races_2018_labelled <- races_2018 |>
  filter(event %in% c("Ultra-Trail Mt. Fuji", 
                      "Le Grand Raid De La Réunion", 
                      "Leadville Trail 100 Run",
                      "Tahoe Rim Trail Endurance Run",
                      "Western States Endurance Run",
                      "HARDROCK 100 ENDURANCE RUN",
                      "6th Elton Volgabus Ultra-Trail®",
                      "La Mision 2018 - VLA",
                      "Hardcore Hundred Miles Trail Ultramarathon"))
```





```{r, fig.height = 6}
ggplot() +
  geom_point(data = races_2018 , aes(x = elevation_gain, y = avg_pace_km)) +
  geom_point(data = races_2018_labelled , aes(x = elevation_gain, y = avg_pace_km), color = "red", size = 3) +
  scale_y_reverse(sec.axis = sec_axis( trans=~.*1.60934, name="Minutes per mile")) +
  labs(title = "\"Very runnable\"",
       subtitle = "Average pace of finishers of 100 mile trail races in 2018") +
  ggrepel::geom_text_repel(data = races_2018_labelled, aes(x = elevation_gain, y = avg_pace_km, label = event), color = "#929299", direction = "y", min.segment.length = 0, hjust= 0) +
  hrbrthemes::theme_ft_rc()
  # geom_point(aes(x = elevation_gain, y = fastest_pace), shape = 1, alpha = 0.3) +
  # geom_segment(aes(x = elevation_gain, xend = elevation_gain, y = fastest_pace, yend= avg_pace))
  
```

## Todo

* Add winning pace
* Right align geom_text_repel labels
* Maybe change theme



```{r}
ggplot() +
  geom_point(data = races_2018 , aes(x = elevation_gain, y = avg_pace_km)) +
  geom_point(data = races_2018_labelled , aes(x = elevation_gain, y = avg_pace_km), color = "red", size = 3) +
  scale_y_reverse(sec.axis = sec_axis( trans=~.*1.60934, name="Minutes per mile")) +
  labs(title = "\"Very runnable\", they said",
       subtitle = "Average pace of finishers of 100 mile trail races in 2018") +
  # ggrepel::geom_text_repel(data = races_2018_labelled, aes(x = elevation_gain, y = avg_pace_km, label = event), color = "#929299", direction = "y", min.segment.length = 0, hjust= 0) +
  geom_point(data = races_2018, aes(x = elevation_gain, y = fastest_pace_km)) +
  geom_segment(data = races_2018, aes(x = elevation_gain, xend = elevation_gain, y = fastest_pace_km, yend= avg_pace_km)) +
  theme_bw()
```

```{r}
bg_color <- "#FEFAE0"
a_highlight_color <- "#BC6C25"
a_light_color <- "#DDA15E"
b_highlight_color <- "#283618"
b_light_color <- "#606C38"

theme_ultra <- theme_bw() +
  theme(plot.background = element_rect(fill = bg_color, color = NA),
        panel.background = element_rect(fill = bg_color, color = NA),
        text = element_text(family = "Roboto Condensed"))
  
```


```{r, fig.height = 6}
ggplot() +
  geom_segment(data = races_2018_labelled, aes(x = elevation_gain, xend = elevation_gain, y = fastest_pace_km, yend= avg_pace_km), alpha = 0.5) +
  geom_segment(data = races_2018, aes(x = elevation_gain, xend = elevation_gain, y = fastest_pace_km, yend= avg_pace_km), alpha = 0.2) +
  geom_point(data = races_2018 , aes(x = elevation_gain, y = avg_pace_km), shape= 21, fill = b_light_color) +
  geom_point(data = races_2018 , aes(x = elevation_gain, y = fastest_pace_km), shape= 21, fill = a_light_color) +
  geom_point(data = races_2018_labelled , aes(x = elevation_gain, y = avg_pace_km), size =3, shape = 21, fill = b_highlight_color) +
  scale_y_reverse(sec.axis = sec_axis( trans=~.*1.60934, name="Minutes per mile")) +
  labs(title = "\"Very runnable\", they said",
       subtitle = "Average pace of finishers and winning pace of 100 mile trail races in 2018") +
  ggrepel::geom_text_repel(data = races_2018_labelled, aes(x = elevation_gain, y = avg_pace_km, label = event),  direction = "y", min.segment.length = 0, hjust= 0) +
  geom_point(data = races_2018_labelled, aes(x = elevation_gain, y = fastest_pace_km), shape = 21, fill = a_highlight_color, size= 3) +
  theme_ultra
```


# DNF - rate

Cannot be calculated, seems not all races listed rank = NA for DNF


```{r}
n_dnf <- ultra_rankings %>% 
  left_join(race) %>% 
  filter(is.na(rank)) %>% 
  group_by(race_year_id) %>% 
  summarise(n_dnf = n())
```


```{r}
race %>% 
  left_join(n_dnf)
```


```{r}
races_2018 <- ultra_rankings |>
  left_join(race) |>
  filter(distance > 155, 
         distance < 170,
         date > as.Date("2017-12-31"),
         date < as.Date("2018-12-31")) |>
  mutate(miles = distance / 1.60934,
         pace_km = (time_in_seconds/60)/distance,
         pace_mile = (time_in_seconds/60)/miles) |>
  group_by(race_year_id) |>
  mutate(avg_pace_mile = mean(pace_mile, na.rm = TRUE),
            avg_pace_km = mean(pace_km, na.rm = TRUE),
            race_year_id = race_year_id,
            fastest_pace_km = min(avg_pace_km),
            participants_from_rankings = n()) |>
  left_join(race) |>
  distinct() |>
  filter(participants_from_rankings > 50) 
  


```








```{r}
sessionInfo()
```


