---
title: "2020-11-10 Historical Phone Usage"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)
library(ggplot2); theme_set(theme_bw())
`%nin%` = Negate(`%in%`)
```

```{r}
mobile <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-11-10/mobile.csv')
landline <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-11-10/landline.csv')
```

## Explore data

```{r}
skim(mobile)
mobile %>% filter(is.na(code))
mobile %>% filter(entity == "Syrian Arab Republic")
mobile %>% filter(is.na(total_pop))

skim(landline)
landline %>% filter(is.na(total_pop))
```
## Combine

```{r}
dat <- mobile %>% select(entity, year, mobile_subs) %>% left_join(landline) # selecting first because NAs are introduced when joining on total_pop and gdp_per_cap
skim(dat)
```

## Potential questions

* What is the relationship between mobile phone subs and landline subs?
* Which countries were first to adopt mobile phone subs?
* Where are the highest rates of change? Fast adoption of mobiles?
* Expectation: Landline down, mobile up - Any examples that contradict that?

## Exploratory plots

* good old pairs()

```{r}
dat %>% select(total_pop, landline_subs, gdp_per_cap, mobile_subs) %>% 
  pairs()
```

* Mobile and Landline subscriptions over time
* The relationship looks similar over the continents, I expect there to be higher differences between the countries

```{r}
dat %>% 
  pivot_longer(cols = c(landline_subs, mobile_subs), names_to = "Subscription", values_to = "SubsPer100") %>% 
  filter(!is.na(continent)) %>% 
  ggplot(aes(x= year, y = SubsPer100, color = Subscription)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~continent) +
  labs(title = "Mobile and Landline subscriptions over time", x = "Year", y = "Subscriptions per 100 inhabitants")
  
```

* Early adopters?
* There seems to be missing data for many years

```{r}
dat %>%
    filter(mobile_subs > 10) %>% 
  pivot_longer(cols = c(landline_subs, mobile_subs), names_to = "Subscription", values_to = "SubsPer100") %>% 
  filter(!is.na(continent), year > 1990, year < 1997) %>% 
  ggplot(aes(x= year, y = SubsPer100, color = Subscription)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~entity) +
  scale_x_continuous(limits = c(1990, 2000), breaks = c(1990, 1992, 1994, 1996, 1998 , 2000)) +
  labs(title = "Early adopters of mobile subscriptions", x = "Year", y = "Subscriptions per 100 inhabitants")
  
```

* At what time do mobile subscriptions overtake landline subscriptions?

```{r}
dat %>% 
  filter(mobile_subs > landline_subs) %>% 
  group_by(entity, continent) %>% 
  summarise(infliction_year = min(year)) %>% 
  ungroup() %>% 
  count(infliction_year) %>% 
  ggplot(aes(x = infliction_year, y = n )) +
  geom_col() +
  labs(title = "At what time do mobile subscriptions overtake landline subscriptions?",
       x = "Year", y = "Number of countries")

```

* Which continents did adopt early or late?

```{r}
infliction <- dat %>% 
  filter(mobile_subs > landline_subs) %>% 
  group_by(entity, continent) %>% 
  summarise(infliction_year = min(year))

country_count <- infliction %>% 
  ungroup() %>% 
  count(infliction_year) 


dat %>% 
  left_join(infliction) %>% 
  left_join(country_count) %>% 
  select(entity, n, infliction_year, continent) %>% 
  distinct() %>% 
  ggplot(aes(x = infliction_year, fill = continent)) +
  geom_bar(stat= "count") +
  labs(title = "At what time do mobile subscriptions overtake landline subscriptions?",
       x = "Year", y = "Number of countries")

infliction %>%
  filter(infliction_year > 2010 | infliction_year < 2000)

```

## Further analysis

* What are the outlying countries? 
* Surely Cambodia is a data entry mistake
* Relation to GDP?
* Look at overall adoption rather than infliction point


