---
title: " 2022-03-01 Alternative Fuel Stations"
author: Florian Tanner
date: "`r format(Sys.time())`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list = ls())

library(tidyverse)
library(ggmap)

library(gganimate)

library(showtext)

sysfonts::font_add_google("Poppins")
showtext_opts(dpi = 300)
showtext_auto(enable = TRUE)

font <- "Poppins"
```

# Read data

```{r}
stations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-03-01/stations.csv') 
```

# Ideas 

* For each state, what is the availability of alternative fuel stations
* Density
* Average distance to stations
* Adjust for population? Adjust for size of the state?

# Notes

* Need to filter for 
  * access_code == "public"

* cng_renewable_source + lng_renewable_source + ev_renewable_source
  * This is interesting: How many stations can do it? 
  
* Vast majority of stations is for electric
  
 group by city or state?

# EDA

```{r}
public_access <- stations |>  
  janitor::clean_names() |> 
  filter(access_code == "public") |> 
  mutate(fuel_type= case_when(fuel_type_code == "ELEC" ~ "Electric",
                              fuel_type_code == "BD" ~ "Biodiesel (B20 and above)",
                              fuel_type_code == "CNG" ~ "Compressed Natural Gas (CNG)",
                              fuel_type_code == "E85" ~ "Ethanol (E85)",
                              fuel_type_code == "HY" ~ "Hydrogen",
                              fuel_type_code == "LNG" ~ "Liquefied Natural Gas (LNG)",
                              fuel_type_code == "LPG" ~ "Propane (LPG)"))
```


How many are planned / expected?

```{r}
table(public_access$status_code)
table(public_access$fuel_type_code)
```


# Maps of individual station

```{r}
us_loc <- c(left = -125, bottom = 25, right = -65, top = 49)
us <- get_stamenmap(bbox= us_loc, source="stamen", maptype="toner-lite", crop=FALSE, zoom = 6)
```


```{r}
p_static <- ggmap(us) + 
  geom_point(data = public_access, aes(x = x, y =y, color = fuel_type_code), size = 0.2)+ 
  ggsci::scale_color_aaas() + 
  facet_wrap(~fuel_type) +
  theme_void() +
  theme(legend.position = "none")
```


```{r}
ggsave(plot = p_static, filename = "test.png", units = "cm", height = 20, width= 20, limitsize = F, dpi = 600 )
```

# Gif

```{r}


p_animate <- ggmap(us) + 
  geom_point(data = public_access, aes(x = x, y =y, color = fuel_type), size = 0.2)+ 
  ggsci::scale_color_aaas() + 
  theme_void() +
  theme(legend.position = "none") +
  transition_states(fuel_type) + 
  ggtitle("{closest_state}") 

p_animation <- animate(p_animate, nframes = 7, fps = 0.5,  height = 12, width = 14, units = "cm", res = 300) 
anim_save("fuels.gif")

```



# Shiny 



```{r}
sessionInfo()
```


















