---
title: "2021-02-09 Wealth and Income"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)
library(ggplot2); theme_set(theme_bw())
library(ggsci)
`%nin%` = Negate(`%in%`)
```

```{r}
income_distribution <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_distribution.csv')
```

```{r}
skim(income_distribution)
```

## Exploring data

* What are the categories for race?
* Checking data completeness -> only "Black Alone" and "White Alone" seem to be complete over the years
* Testing that income_distribution adds up to 100 for a single year and race?

```{r}
income_distribution %>% 
  count(race)

income_distribution %>% 
  filter(race == "Asian Alone", year == 2000) %>% 
  summarise(sum_in_dist = sum(income_distribution))
```

* Data seems to be complete after 2001 (maybe the categories for race changed?)

```{r}
income_distribution %>% 
  filter(year > 2001) %>% 
  count(race)
```

* Plotting number of households over time per race

```{r}
income_distribution %>% 
  ggplot(aes(x = year, y = number / 1000, color = race)) +
  geom_smooth(se = F)
```



* Graphing income distribution over the years
* It seems the distribution is tending up over time for all races
* Try a barplot

```{r}
correct_order_income_bracket <- unique(income_distribution$income_bracket)

income_distribution %>% 
  mutate(income_bracket = fct_relevel(income_bracket, correct_order_income_bracket)) %>% 
  ggplot(aes(x = income_bracket, y = income_distribution, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle=45, hjust = 1)) +
  facet_wrap(~race) +
  scale_color_viridis_c()
```
* 

```{r}
income_distribution %>% 
  filter(year %in% c(1970, 1980, 1990, 2000, 2010, 2019)) %>% 
  mutate(income_bracket = fct_relevel(income_bracket, rev(correct_order_income_bracket))) %>% 
  ggplot(aes(x = race, y = income_distribution, color = income_bracket, fill = income_bracket)) +
  geom_col(width= 0.8) +
  theme(axis.text.x = element_text(angle=45, hjust = 1)) +
  facet_wrap(~year) +
  scale_color_jco() +
  scale_fill_jco() 
```

* Summarise over centuries, tidying up graph

```{r}
income_distribution$century <- "2010s"
income_distribution$century[income_distribution$year < 2010] <- "2000s"
income_distribution$century[income_distribution$year < 2000] <- "1990s"
income_distribution$century[income_distribution$year < 1990] <- "1980s"
income_distribution$century[income_distribution$year < 1980] <- "1970s"
income_distribution$century[income_distribution$year < 1970] <- "1960s"

income_distribution %>% 
  group_by(race, century, income_bracket) %>% 
  summarise(per_century_inc_dist = mean(income_distribution)) %>% 
  mutate(income_bracket = fct_relevel(income_bracket, rev(correct_order_income_bracket))) %>% 
  ggplot(aes(x = race, y = per_century_inc_dist, color = income_bracket, fill = income_bracket)) +
  geom_col(width= 0.8) +
  theme(axis.text.x = element_text(angle=45, hjust = 1)) +
  facet_wrap(~century) +
  scale_color_jco() +
  scale_fill_jco() + 
  labs(title = "Income distribution per race over the last four centuries",
       x = "",
       y = "Income distribution",
       color= "Income bracket",
       fill = "Income bracket")
```

* Repeat the same graph, just reverse order of x-axis and facets
```{r}
income_distribution %>% 
  group_by(race, century, income_bracket) %>% 
  summarise(per_century_inc_dist = mean(income_distribution)) %>% 
  mutate(income_bracket = fct_relevel(income_bracket, rev(correct_order_income_bracket))) %>% 
  ggplot(aes(x = century, y = per_century_inc_dist, color = income_bracket, fill = income_bracket)) +
  geom_col(width= 0.8) +
  theme(axis.text.x = element_text(angle=45, hjust = 1)) +
  facet_wrap(~race) +
  scale_color_jco() +
  scale_fill_jco() + 
  labs(title = "Income distribution per race over the last four centuries",
       x = "",
       y = "Income distribution",
       color= "Income bracket",
       fill = "Income bracket")
```

## Final summary plot 

* With this type of plot, there is no need to summarise over centuries
* Limit plot to four race categories to make trends visible


```{r}
income_distribution %>% 
  filter(race %in% c("Black Alone", "White Alone, Not Hispanic", "Hispanic (Any Race)", "Asian Alone")) %>%
  mutate(income_bracket = fct_relevel(income_bracket, rev(correct_order_income_bracket))) %>% 
  ggplot(aes(x = year, y = income_distribution, color = income_bracket, fill = income_bracket)) +
  geom_col(width= 0.8) +
  facet_wrap(~race) +
  scale_color_jco() +
  scale_fill_jco() +
  labs(title = "US household income per race ",
       x = "Year",
       y = "Income distribution",
       color= "Income bracket",
       fill = "Income bracket")
```

```{r}
sessionInfo()
```

