---
title: "2021-01-26 	Plastic Pollution"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)
library(ggplot2); theme_set(theme_bw())
`%nin%` = Negate(`%in%`)
```


```{r}
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv')
skim(plastics)
```

## Tidying format

Missing data: 

* What sort of plastic is not always recorded, will stick to looking at grand_total
* Country seems to be missing quite often, will remove "EMPTY"

```{r}
plastics_long <- plastics %>%
  filter(country != "EMPTY") %>% 
  pivot_longer(cols = c("hdpe", "ldpe", "o", "pet", "pp", "ps", "pvc", "grand_total"),
               names_to = "type",
               values_to= "amount") %>% 
  select(-empty)

plastics_long %>% 
  count(parent_company) %>% 
  arrange(desc(n))

```
## Possible questions

* Activity of BFFP per country?
* Growth of activity over two years?
* Which companies pollute most?
* What sort of plastics?

```{r}
plastics_long %>% 
  select(country, year, num_events) %>% 
  distinct() %>% 
  arrange(desc(num_events))

```
## Most activity of BFFP in 2020

```{r}

plastics_long %>% 
  group_by(country, year) 

plastics_long %>% 
  select(year, country, volunteers) %>% 
  filter(country != "EMPTY") %>%
  distinct() %>% 
  group_by(country) %>% 
  mutate(total_volunteers = sum(volunteers)) %>% 
  ungroup %>% 
  mutate(country = fct_reorder(country, total_volunteers)) %>%
  arrange(desc(total_volunteers)) %>% 
  head(25) %>% 
  ggplot(aes(x = volunteers, y= country, fill = as.factor(year))) +
  geom_col(position = position_dodge2())
```

## Has there been less reporting or less volunteering in 2020? Or has the participation changed that much?

* Hard to tell
* The change between the years is nonetheless interesting

```{r}
plastics_long %>% 
  select(year, country, volunteers) %>% 
  filter(country != "EMPTY") %>%
  distinct() %>% 
  group_by(country) %>% 
  mutate(total_volunteers = sum(volunteers)) %>% 
  ungroup %>% 
  mutate(country = fct_reorder(country, total_volunteers)) %>%
  arrange(desc(total_volunteers)) %>% 
  head(50) %>% 
  ggplot(aes(x = volunteers, y= country)) +
  geom_col(position = position_dodge2()) +
  facet_wrap(~year, nrow = 1)
```

## Which company produced the most waste overall?

```{r}
plastics_long %>% 
  mutate_at(vars(parent_company), function(x){gsub('[^ -~]', '', x)}) %>%  # removing non UTF-8 character that throw off summarise()
  filter(type == "grand_total", 
         year == "2019",
         parent_company %nin% c("Grand Total",
                                "Unbranded",
                                "Assorted")) %>% # Going through these iteratively to remove irrelevant parent companies
  group_by(parent_company) %>% 
  summarise(total_per_company = sum(amount)) %>% 
  arrange(desc(total_per_company)) %>% 
  head(20) %>% 
  mutate(parent_company = fct_reorder(parent_company, total_per_company)) %>% 
  ggplot(aes(x = total_per_company, y = parent_company)) +
  geom_col()

```

## Further analysis

* Normalize for participation per country?
* Companies x Types of plastics?

```{r}
sessionInfo()
```


