---
title: "2021-06-29 Animal Rescues"
output: github_document
---

```{r setup, include=FALSE}
.libPaths("C:/R-packages2/")
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r}
library(tidyverse)
library(ggmap)
theme_set(theme_bw())
```

## Read data 

```{r}
animal_rescues <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-29/animal_rescues.csv')
```

## EDA

What animals are involved?

* Cats are most common
* 12 incidents with snakes - what happened?
* 2 incidents with fish - what happened?

```{r}
animal_rescues %>% 
  group_by(animal_group_parent) %>% 
  count
```

### Looking at a few incidents

FISH IN DANGER OF DYING IN POND - more details please :) 

```{r}
animal_rescues %>% 
 filter(animal_group_parent %in% c("Horse"))
```

### Animals rescued from height
* Apparently a snake was rescued from height, what happened there? 
* Will focus on cats on trees, where does that happen?

```{r}
animal_rescues %>% 
  filter(special_service_type_category == "Animal rescue from height") %>% 
  group_by(animal_group_parent) %>% 
  count
```


```{r}
cats <- animal_rescues %>% 
  filter(special_service_type_category == "Animal rescue from height",
         animal_group_parent %in% c("cat", "Cat")) %>% 
  mutate(latitude = as.numeric(latitude), longitude= as.numeric(longitude)) %>% 
  filter(!is.na(latitude))
```

```{r}
horses <- animal_rescues %>% 
  filter(special_service_type_category == "Animal rescue from water",
         animal_group_parent %in% c("Horse")) %>% 
  mutate(latitude = as.numeric(latitude), longitude= as.numeric(longitude)) %>% 
  filter(!is.na(latitude))
```

## ggmap approach

```{r}
summary(as.numeric(animal_rescues$longitude))
summary(as.numeric(animal_rescues$latitude))
```


```{r}
london_loc <- c(bottom = 51.20000, top = 51.62000, left = -0.58000, right = 0.37000)
london <- get_stamenmap(bbox= london_loc, source="stamen", maptype="terrain", crop=FALSE)
```


```{r}
p_lond <- ggmap(london) +
  geom_point(data = horses, 
             aes(x = longitude , y = latitude),
             alpha = 0.5,
             color = "blue",
             size = 1) +
  theme_void() +
  labs(title = "52 horses were rescued from water by the\nLondon fire brigade in the last 12 years",
       caption = "Data: London.gov | Map: Stamen | Graphic: @TannerFlorian") +
  theme(legend.position = "NULL",
        plot.title = element_text(size = 6, hjust = 0.5, vjust = 2),
        plot.caption = element_text(size = 4.5, hjust = 1))
  
```

```{r}
ggsave(plot = p_lond, filename = "horses.png", device = "png",  units = "in", width = 4, height = 2.25, scale = 1, type = "cairo", dpi = 600)
```

```{r}
sessionInfo()
```


