---
title: "2021-02-23 Employment and Earnings"
output: github_document
---

```{r setup, include=FALSE}
.libPaths("C:/R-packages2/")
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
library(skimr)
library(ggsci)
theme_set(theme_bw())
`%nin%` = Negate(`%in%`)
```


```{r}
employed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-23/employed.csv')
earn <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-23/earn.csv')

```
## First look

* Will focus on 'employed' dataset
* Possible questions:
  * What are womens main industries per year (per race)? 
  * What are mens main industries per year (per race)?

```{r}
skim(earn)
skim(employed)
```

## 'Employed' exploratory

* If 'industry' is used, a couple of things probably need to be fixed:
  * remove NA, fix parsing error for 'mining, quarrying...'
  * Combine or remove categories that count numbers twice



```{r}
employed %>% 
  count(industry) %>% 
  arrange(n)
```

```{r}
employed %>% 
  count(minor_occupation) %>% 
  arrange(n)
```
* that might mean that it is not possible to answer questions per race per gender
* maybe just analyze per race instead?

```{r}
employed %>% 
  count(race_gender) %>% 
  arrange(n)
```

### Data prep

* Fixing parsing errors
* pivot_wider for easy ratio calculations


```{r}
employed_clean <- employed
employed_clean$minor_occupation[employed_clean$minor_occupation == "Manage-ment, business, and financial operations occupations"] <- 
  "Management, business, and financial operations occupations"

employed_clean$industry[employed_clean$industry == "Mining, quarrying, and\r\noil and gas extraction"] <- 
  "Mining, quarrying, and oil and gas extraction"

employed_wide <- employed_clean  %>% 
  pivot_wider(id_cols = c(year, minor_occupation, major_occupation, industry),  names_from = race_gender, values_from = employ_n) %>% 
  mutate(r_asian = Asian/TOTAL, 
         r_black_or_afram = `Black or African American`/TOTAL, 
         r_white = White/TOTAL,
         r_male = Men/TOTAL, 
         r_female = Women/TOTAL)


```

* Is it possible to find out about minor occupation over all industries? 
  * Ideally, this should be normalized for the total size of the industries

```{r}
per_minor_occ <- employed_wide %>% 
  group_by(year, minor_occupation) %>% 
  summarize(mean_r_white = mean(r_white, na.rm = T), 
            mean_r_asian = mean(r_asian, na.rm = T),
            mean_r_boaa = mean(r_black_or_afram, na.rm = T),
            mean_r_male = mean(r_male, na.rm =T),
            mean_r_female = mean(r_female, na.rm = T)) %>% 
  pivot_longer(cols = c(-year, -minor_occupation), names_to = "ratio")

```

### 'employed' graphs

* Graph of ratio of race per minor_occupation in 2015
* there is missing data (other category for race?)  + no striking differences
  * the only other categories for race_gender in this dataset are "Men", "Women", "Total"

```{r}
per_minor_occ %>% 
  filter(year == 2015, ratio %in% c("mean_r_asian", "mean_r_white", "mean_r_boaa")) %>% 
  ggplot(aes(x = value, y = minor_occupation, color = ratio, fill = ratio)) +
  geom_col()
```


* Graph of ratio of gender per minor_occupation in 2015
  * Some more striking differences!


```{r}
per_minor_occ %>% 
  filter(year == 2015, ratio %in% c("mean_r_female", "mean_r_male")) %>% 
  ggplot(aes(x = value, y = minor_occupation, color = ratio, fill = ratio)) +
  geom_col()
```

* Now to expand this graph over the years
  * the ratios of men and women do not add up to 1 but there are no other gender categories


```{r}
per_minor_occ %>% 
  filter(ratio %in% c("mean_r_female", "mean_r_male")) %>%
  ggplot(aes(x = year, y = value, color = ratio, fill = ratio)) +
  geom_col() +
  facet_wrap(~minor_occupation)+ 
  labs(title = "Ratio of gender in minor occupations, averaged over all industries") +
  theme(legend.position = "bottom")
```

### Big summary plot

* It does not really need the resolution over time, nothing drastic happens
  * take average over last 5 years 

```{r}
per_minor_occ$category <- "Race"
per_minor_occ$category[per_minor_occ$ratio == "mean_r_female" | per_minor_occ$ratio == "mean_r_male" ] <- "Gender"

per_minor_occ %>% 
  ggplot(aes(x = year, y = value, color = ratio, fill = ratio)) +
  geom_line() +
  facet_grid(category~minor_occupation)+ 
  labs(title = "Ratio of gender in minor occupations, averaged over all industries") +
  theme(legend.position = "bottom")
```

* It does not really need the resolution over time, nothing drastic happens
  * take average over last 5 years 
  * Rename subgroups
* Sort y-axis alphabetically

  

```{r}
per_minor_occ$category <- "Race"
per_minor_occ$category[per_minor_occ$ratio == "mean_r_female" | per_minor_occ$ratio == "mean_r_male" ] <- "Gender"

per_minor_occ$subgroup[per_minor_occ$ratio == "mean_r_female"] <- "Women"
per_minor_occ$subgroup[per_minor_occ$ratio == "mean_r_male"] <- "Men"
per_minor_occ$subgroup[per_minor_occ$ratio == "mean_r_asian"] <- "Asian"
per_minor_occ$subgroup[per_minor_occ$ratio == "mean_r_boaa"] <- "Black or African American"
per_minor_occ$subgroup[per_minor_occ$ratio == "mean_r_white"] <- "White"

p_employed_minor_occ_summary <- per_minor_occ %>% 
  group_by(minor_occupation, ratio, category, subgroup) %>%
  mutate(minor_occupation = str_remove(minor_occupation, " occupations")) %>% 
  summarize(avg_ratio = mean(value)) %>% 
  ggplot(aes(x = avg_ratio, y = fct_rev(minor_occupation), color = subgroup, fill = subgroup)) +
  geom_col(width = 0.7) +
  facet_wrap(~category)+ 
  labs(title = "Ratio of race and gender in minor occupations",
       subtitle ="Averaged over all industries from 2015-2020", 
       color = NULL, 
       fill = NULL,
       y = NULL,
       x = "Ratio") +
  theme(legend.position = "bottom",
        axis.ticks.y = element_blank()) +
  scale_fill_jco() +
  scale_color_jco()

p_employed_minor_occ_summary 
```


```{r}
ggsave(p_employed_minor_occ_summary, filename = "p_employed_minor_occ_summary.png", units = "cm", width = 14, height = 10, limitsize = F, scale = 1.4)
```

## Thoughts

* Deal with missing data
* Should add importance of the minor occupations (how many people overall work in these minor occupations)


```{r}
sessionInfo()
```


















