---
title: "2021-07-20 US Droughts"
output: github_document
---

```{r setup, include=FALSE}
.libPaths("C:/R-packages2/")
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r, message= FALSE, marning = FALSE, echo = FALSE}
library(tidyverse)
library(usmap)
extrafont::loadfonts(device = "win", quiet = TRUE)

`%nin%` <- Negate(`%in%`)
```

## Read data 

```{r}
drought <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-20/drought.csv')
```

## EDA 
* Percentage of state in extreme or exceptional drought
* Use last reported week 
* D3 

Potential deeper analysis
* Compare this to other years
* Average over seasons/ months
  * This data only extends to 2001, so comparisons are difficult


```{r}
drought %>% 
  filter(valid_start == "2021-07-13",
         state_abb == "CA")
```

```{r}
drought_summary <- drought %>% 
  filter(valid_start == "2021-07-13",
         drought_lvl %in% c("D3", "D4")) %>% 
  pivot_wider(id_cols = state_abb, names_from = drought_lvl, values_from = area_pct) %>% 
  mutate(drought = as.integer(D4 + (D3 - D4)),
         fips = fips(state_abb)) %>% 
  select(fips, drought)
```



```{r}
scale_color_high <- "#b30000"
scale_color_low <- "#e6e6e6"
credits <- "The U.S. Drought Monitor is jointly produced by the National Drought Mitigation Center at the University of Nebraska-Lincoln, the United States Department of Agriculture, and the National Oceanic and Atmospheric Administration."

 
p_map <- plot_usmap(data = drought_summary, values = "drought", color = "#747474",
           labels = TRUE) + 
  labs(title = "Drought in mid-July 2021",
       subtitle = "Percentage of area in extreme drought or exceptional drought, the two most severe categories",
       caption = paste0(str_wrap(credits), "\n\nGraphic: @TannerFlorian")) + 
  scale_fill_continuous(low = scale_color_low, high = scale_color_high, 
                         name = "", limits = c(0,100)) +
  theme(panel.background=element_blank(),
        legend.position = "right")

p_map$layers[[2]]$aes_params$size <- 2
p_map
```

```{r}
ggsave(p_map, filename = "drought_map.png", device = "png", units = "cm", width = 14, height = 12, scale = 1, type = "cairo")
```


### Credit

The U.S. Drought Monitor is jointly produced by the National Drought Mitigation Center at the University of Nebraska-Lincoln, the United States Department of Agriculture, and the National Oceanic and Atmospheric Administration. Map courtesy of NDMC.


```{r}
sessionInfo()
```


